Comparing Models of Delta-Notch Signalling

### 2025-04-11 paper/paper.tex
Capitalise delta in Table1; rewrite key assumptions in Appendix C

\begin{table}[!htp]
\centering
\begin{tabular}{|m{8em}|m{5em}|m{20em}|} 
 \hline
 Reaction & Rate & Description \\ 
 \hline
 $N_{A} + D_{B} \rightarrow I_{A}$ & 
 $k_{T} N_{A}D_{B}$ &
 A Notch receptor from cell $A$ binds to a delta ligand from cell $B$, triggering the release of NICD in cell $A$. \\
 \hline
 $N_{A} \rightarrow \emptyset$ & 
 $\gamma N_{A}$ & 
 A Notch receptor from cell $A$ decays. \\
 \hline
 $D_{A} \rightarrow \emptyset$ & 
 $\gamma D_{A}$ & 
 A Delta receptor from cell $A$ decays. \\
 \hline
 $I_{A} \rightarrow \emptyset$ &
 $\gamma_{I} I_{A}$ &
 A NICD molecule from cell $A$ decays.  \\
 \hline
 $\emptyset \rightarrow N_{A}$ & 
 $H^{+}(I_{A})$ &
 A Notch receptor in cell $A$ is produced \\
 \hline
 $\emptyset \rightarrow D_{A}$ &
 $H^{-}(I_{A})$ &
 A Delta ligand in cell $A$ is produced \\
 \hline
\end{tabular}
\caption{
  A list of possible reactions and their rates in a cell $A$ with neighbour $B$.
  Since there are $6$ possible reactions per cell, a system with $n$ cells will have $6n$ possible reactions.
  In the Reaction column, $N_A$, $D_B$, and $I_A$ represent one Notch, Delta, and NICD, respectively. In the Rate column, $N_A$, $D_B$, and $I_A$ refer to the number of Notch molecules in cell A, Delta in cell B, and NICD in cell A, respectively.
}
\label{tb:reactions}
\end{table}

@@

\subsection*{Key Assumptions}

Throughout this derivation, we make the following assumptions:

\begin{enumerate}
  \item The standard deviation of the molecular counts is small compared to their mean values. Specifically:
  \begin{itemize}
    \item $\sqrt{\text{Var}(N)} \ll \langle n \rangle$
    \item $\sqrt{\text{Var}(D)} \ll \langle d \rangle$
    \item $\sqrt{\text{Var}(I)} \ll \langle i \rangle$
  \end{itemize}

  \item As a consequence of assumption 1, we can approximate:
  \begin{itemize}
    \item $\langle n^2 \rangle \approx \langle n \rangle^2$
    \item $\langle d^2 \rangle \approx \langle d \rangle^2$
    \item $\langle i^2 \rangle \approx \langle i \rangle^2$
  \end{itemize}

  \item The molecular counts are approximately independent, allowing us to factorize expected values:
  \begin{itemize}
    \item $\langle n \cdot d \rangle \approx \langle n \rangle \cdot \langle d \rangle$
    \item $\langle n \cdot i \rangle \approx \langle n \rangle \cdot \langle i \rangle$
    \item $\langle d \cdot i \rangle \approx \langle d \rangle \cdot \langle i \rangle$
  \end{itemize}

  \item For functions $H^+$ and $H^-$, we approximate:
  \begin{itemize}
    \item $\langle H^+(i) \rangle \approx H^+(\langle i \rangle)$
    \item $\langle H^-(i) \rangle \approx H^-(\langle i \rangle)$
    \item $\langle n \cdot H^+(i) \rangle \approx \langle n \rangle \cdot H^+(\langle i \rangle)$
    \item $\langle d \cdot H^-(i) \rangle \approx \langle d \rangle \cdot H^-(\langle i \rangle)$
  \end{itemize}
\end{enumerate}

>>>>

\begin{table}[!htp]
\centering
\begin{tabular}{|m{8em}|m{5em}|m{20em}|} 
 \hline
 Reaction & Rate & Description \\ 
 \hline
 $N_{A} + D_{B} \rightarrow I_{A}$ & 
 $k_{T} N_{A}D_{B}$ &
 A Notch receptor from cell $A$ binds to a Delta ligand from cell $B$, triggering the release of NICD in cell $A$. \\
 \hline
 $N_{A} \rightarrow \emptyset$ & 
 $\gamma N_{A}$ & 
 A Notch receptor from cell $A$ decays. \\
 \hline
 $D_{A} \rightarrow \emptyset$ & 
 $\gamma D_{A}$ & 
 A Delta receptor from cell $A$ decays. \\
 \hline
 $I_{A} \rightarrow \emptyset$ &
 $\gamma_{I} I_{A}$ &
 A NICD molecule from cell $A$ decays.  \\
 \hline
 $\emptyset \rightarrow N_{A}$ & 
 $H^{+}(I_{A})$ &
 A Notch receptor in cell $A$ is produced \\
 \hline
 $\emptyset \rightarrow D_{A}$ &
 $H^{-}(I_{A})$ &
 A Delta ligand in cell $A$ is produced \\
 \hline
\end{tabular}
\caption{
  A list of possible reactions and their rates in a cell $A$ with neighbour $B$.
  Since there are $6$ possible reactions per cell, a system with $n$ cells will have $6n$ possible reactions.
  In the Reaction column, $N_A$, $D_B$, and $I_A$ represent one Notch, Delta, and NICD, respectively. In the Rate column, $N_A$, $D_B$, and $I_A$ refer to the number of Notch molecules in cell A, Delta in cell B, and NICD in cell A, respectively.
}
\label{tb:reactions}
\end{table}

@@

\subsection*{Key Assumptions}
Throughout this derivation, we make the following assumptions:
\begin{enumerate}
  \item Fluctuations in molecular counts are small relative to their mean values; that is, $\sqrt{\text{Var}(N)} \ll \langle n \rangle$, $\sqrt{\text{Var}(D)} \ll \langle d \rangle$, and $\sqrt{\text{Var}(I)} \ll \langle i \rangle$.

  \item Second moments can be approximated by the squares of the means, i.e., $\langle n^2 \rangle \approx \langle n \rangle^2$, $\langle d^2 \rangle \approx \langle d \rangle^2$, and $\langle i^2 \rangle \approx \langle i \rangle^2$.

  \item Molecular species are approximately uncorrelated, so that joint moments factorize, e.g., $\langle n \cdot d \rangle \approx \langle n \rangle \langle d \rangle$.

  \item For Hill functions $H^+$ and $H^-$, we approximate expectation values by evaluating the function at the mean, i.e., $\langle H(i) \rangle \approx H(\langle i \rangle)$, and similarly $\langle x \cdot H(i) \rangle \approx \langle x \rangle H(\langle i \rangle)$.
\end{enumerate}
